RNAseq aCRY pCRY data dive (separate)

### 2024-04-26 2_A_dive_aCRY_pCRY (combined)/2_A_dive_aCRY_pCRY.Rmd
acry pcry separate

# Load System
## -R_libraries
```{r r_libraries}
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

@@

library(viridis)
library(patchwork)
library(ggpubr)
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)

@@

# Seq 1
tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm"
load(file=tximeta_pcry_file)
gse_pcry <- gse

@@

# colData(dds) %>% head() %>% kable()
# add metadata
# colData(dds)$strain <- factor(colData(dds)$strain)

# ## combine technical replicates #####\n
# colData(dds)
# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
# colData(dds)
# dds$runsCollapsed
# colnames(dds)
# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))

@@

# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

@@

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
ntd <- normTransform(dds)
rld <- rlog(dds, blind=FALSE) #regularized logarithm
```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
# load(file=paste(data,"rlog.rld", sep="/"))
meanSdPlot(counts(dds, normalized =TRUE))
meanSdPlot(assay(ntd))
meanSdPlot(assay(vsd))
meanSdPlot(assay(rld))

@@

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                    genotype = c("green4","orchid","tan2"))

names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

@@

@@

# plot Scree
# find explaining PCs
horn <- parallelPCA(vst_dat)
# horn$n = 2
# warnings()

# horn <- list()
# horn$n <- 2

@@

# 3. Results
## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}

@@

## Plot Counts
```{r}

## Volcanos
### dark
```{r}

@@

### blue
```{r}

@@

>>>>

# Load System
## -R_libraries
```{r r_libraries}
library(patchwork)
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

@@

library(viridis)

library(ggpubr)
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)

@@

# Seq 1
ifelse(Sys.info()["sysname"]== "Linux",
       tximeta_pcry_file <- dir %>% dirname() %>% paste(.,"P3044/tximeta.txm",sep="/"),
    ifelse(Sys.info()["sysname"]== "Darwin",
       tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm",
       tximeta_pcry_file <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044/tximeta.txm"))

load(file=tximeta_pcry_file)
gse_pcry <- gse

@@

# remove outlier
# outliner <- c("WTdark1")
# colData(dds)[outliner,]
# keep <- colnames(dds) %>% .[!str_detect(.,outliner)]
# dds <- dds[,keep]

@@

# save(dds, file = paste(dir,"dds_acry_pcry_all.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

@@

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
ntd <- normTransform(dds)
# rld <- rlog(dds, blind=FALSE) #regularized logarithm
# save(rld, file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"), compress = FALSE)
```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
load(file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"))
meanSdPlot(counts(dds, normalized =TRUE))
meanSdPlot(assay(ntd))
meanSdPlot(assay(vsd))
meanSdPlot(assay(rld))

# (m1+m2) / (m3)
# patchwork::wrap_plots(m1)

@@

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
                    genotype = c("green4","orchid","tan2"))
#                    "ivory","yellow1","gold"
names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

@@

# -> how to handel batch effects: https://www.biostars.org/p/266507/

@@

# plot Scree
# find explaining PCs
# horn <- parallelPCA(vst_dat)
# horn$n = 5 # aCRY&pCRY
# warnings()
# 
horn <- list()
horn$n <- 5

@@

# 3. Results
## Colors
```{r}
# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
# group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")

anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                    genotype = c("green4","orchid","tan2"))

group.colours2 <- c("#00B500","#007E56","#805B00","#FF00AB","#8024AB","#FF0056","tan2","brown","orangered")
```

## Plot Counts
```{r plot_counts}

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts1

gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = anno_colors$genotype) +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts1 / gcounts2

# ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
# width = 12,
# height = 8)

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

# Plot
max_val <- 1.0*max(all_counts$count)
gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts1

gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = anno_colors$genotype) +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts3 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colours2) +
  scale_y_continuous(trans = "log2")
gcounts3

gcounts1 / gcounts2 / gcounts3

# ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
#   width = 12,
#   height = 8)

## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}

design(dds)
resultsNames(dds)
cbind(resultsNames(dds),c(rep(0,resultsNames(dds) %>% length())))

@@

## Volcanos
### dark
```{r, eval=FALSE}

@@

### blue
```{r eval=FALSE}

@@

## Volcanos
### dark
```{r, eval=FALSE}

### blue
```{r, eval=FALSE}

### red
```{r, eval=FALSE}

### all
```{r, eval=FALSE}

### 2024-04-26 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
acry pcry separate

---
title: "RNAseq aCRY pCRY data dive"
author: "Kelterborn"
date: "2024-04-19"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'Readme.md')
        })
---

@@

# Seq 1
ifelse(Sys.info()["sysname"]== "Linux",
       tximeta_pcry_file <- dir %>% dirname() %>% paste(.,"P3044/tximeta.txm",sep="/"),
    ifelse(Sys.info()["sysname"]== "Darwin",
       tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm",
       tximeta_pcry_file <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044/tximeta.txm"))

load(file=tximeta_pcry_file)
gse_pcry <- gse

@@

# combine
( mcols(gse_acry) == mcols(gse_pcry) ) %>% summary()
# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
names(colData(gse_pcry)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
colData(gse_acry) <- colData(gse_acry)[,c(1:4,7:12)]
colData(gse_pcry)$experiment <- "pcry"
colData(gse_pcry) <- colData(gse_pcry)[,c(1,3,2,4:6,8,12,7,11)]  
# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()

# now same metadata
gse <- cbind(gse_acry,gse_pcry)

colData(gse)$genotype

colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","pcry_dark","WT_blue","acry_blue","pcry_blue","WT_red","acry_red","pcry_red"))

colData(gse)$experiment

@@

## Run Deseq2
```{r dds_acry_pcry}
# design <- ~condition
design <- ~genotype+treatment+genotype:treatment

dds <- DESeqDataSet(gse, design=design)
colData(dds) %>% head() %>% kable()

@@

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$experiment)

head(mcols(dds)$geneSymbol)

@@

# save(dds, file = paste(dir,"dds_acry_pcry_all.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

@@

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
ntd <- normTransform(dds)
# rld <- rlog(dds, blind=FALSE) #regularized logarithm
# save(rld, file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"), compress = FALSE)
```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
load(file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"))
meanSdPlot(counts(dds, normalized =TRUE))
meanSdPlot(assay(ntd))
meanSdPlot(assay(vsd))
meanSdPlot(assay(rld))

# (m1+m2) / (m3)
# patchwork::wrap_plots(m1)

@@

### - Check sample distance
```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
                    genotype = c("green4","orchid","tan2"))
#                    "ivory","yellow1","gold"
names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         annotation_col=anno_col,
         annotation_colors = anno_colors,
         show_colnames     = FALSE,
         col=viridis(20),
         cutree_rows = 6,
         cutree_cols = 6,
         fontsize_row = 5)

# -> how to handel batch effects: https://www.biostars.org/p/266507/

@@

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))
vst_dat <- assay(vst(dds))

@@

g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "experiment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = anno_colors$genotype)

g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "condition") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  # coord_fixed()
  scale_color_viridis_d(option ="viridis")

((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")

@@

###### -- Advanced PCA
```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# calculate PCA (all data)
## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots

# p <- pca(vst_dat, metadata = colData(dds))
# vst <- assay(vst(dds))
p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)

# check different PCAs
p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment

p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype

p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition

(p1+p2) / (p4)

# plot Scree
# find explaining PCs
# horn <- parallelPCA(vst_dat)
# horn$n = 5 # aCRY&pCRY
# warnings()
# 
horn <- list()
horn$n <- 5

screeplot(p,
    components = getComponents(p, 1:20),
    vline = c(horn$n, elbow)) +
    geom_label(aes(x = horn$n + 1, y = 30,
      label = paste('Horn',horn$n,sep="="), vjust = -1, size = 8)) +
    geom_label(aes(x = elbow + 1, y = 50,
      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8))

bi <- biplot(p,x="PC3",y="PC1",
    lab = p$metadata$condition,
    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    labSize = 3, legendIconSize = 4.0,
    legendPosition = 'bottom',
    sizeLoadingsNames = 3,
    axisLabSize = 10,
    captionLabSize = 1)

@@

layout <- c(
  area(t = 1, l = 1, b = 6, r = 6),
  area(t = 5, l = 1, b = 7, r = 3)
)
pairs + bi + 
  plot_layout(design = layout)

@@

plotloadings(p,
             components = getComponents(p)[1:7],
             col = viridis(3),
             labSize = 3)

@@

## Plot Counts
```{r plot_counts}

@@

# Plot
gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts1

gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = anno_colors$genotype) +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts1 / gcounts2

@@

# Plot
max_val <- 1.0*max(all_counts$count)
gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts1

gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = anno_colors$genotype) +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts3 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colours2) +
  scale_y_continuous(trans = "log2")
gcounts3

gcounts1 / gcounts2 / gcounts3

@@

@@

### red
```{r}

@@

### all
```{r}

>>>>

---
title: "RNAseq aCRY pCRY data dive (separate)"
author: "Kelterborn"
date: "2024-04-19"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'Readme.md')
        })
---

@@

# Seq 1 (pcry)
ifelse(Sys.info()["sysname"]== "Linux",
       tximeta_pcry_file <- dir %>% dirname() %>% paste(.,"P3044/tximeta.txm",sep="/"),
    ifelse(Sys.info()["sysname"]== "Darwin",
       tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm",
       tximeta_pcry_file <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044/tximeta.txm"))

load(file=tximeta_pcry_file)

colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","pcry_dark","WT_blue","pcry_blue","WT_red","pcry_red"))

names(colData(gse)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
colData(gse)$experiment <- "pcry"
gse_pcry <- gse

@@

colData(gse_acry)$treatment <- colData(gse_acry)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
  str_replace(pattern="BL", replacement = "blue") %>%
  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))

colData(gse_acry)$condition <- paste(colData(gse_acry)$genotype,colData(gse_acry)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red"))

@@

## Run Deseq2
```{r dds_acry_pcry}

# design <- ~condition
exp <- c("acry","pcry")
gse_list <- list(gse_acry,gse_pcry)
names(gse_list) <- exp
dds_list <- list()
i <- 1
for (e in names(gse_list)){
 gse <- gse_list[[e]]
 design <- ~genotype+treatment+genotype:treatment
dds <- DESeqDataSet(gse, design=design)
colData(dds) %>% head() %>% kable()

@@

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$condition)

assign(paste("dds",e,sep="_"),dds)
dds_list[[e]] <- dds
}

@@

# save(dds_list, file = paste(dir,"dds_list.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

@@

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}

vsd_list <- lapply(dds_list,vst,blind=FALSE)
ntd_list <- lapply(dds_list,normTransform)
rld_list <- lapply(dds_list,rlog, blind=FALSE)
# save(rld_list, file = paste(dir,"rlog_list.RDS",sep="/"), compress = FALSE)

@@

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
# load(file = paste(dir,"rlog_list.RDS",sep="/"))
msdp1 <- lapply(lapply(dds_list,assay,normalized =TRUE),meanSdPlot)
msdp2 <- lapply(lapply(vsd_list,assay,normalized =TRUE),meanSdPlot)
msdp3 <- lapply(lapply(ntd_list,assay,normalized =TRUE),meanSdPlot)
msdp4 <- lapply(lapply(rld_list,assay,normalized =TRUE),meanSdPlot)

@@

### - Check sample distance
```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
vsd <- vsd_list[[1]]
for (vsd in vsd_list){
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
                    genotype = c("green4","orchid"))
names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         annotation_col=anno_col,
         annotation_colors = anno_colors,
         show_colnames     = FALSE,
         col=viridis(20),
         cutree_rows = 6,
         cutree_cols = 6,
         fontsize_row = 5) %>% print()
}
# -> how to handel batch effects: https://www.biostars.org/p/266507/

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))

for (vsd in vsd_list){

@@

g2 <- ggplot(pcaData, aes(PC1, PC2, color=genotype, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "genotype") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = anno_colors$genotype)

g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype, label=clientName)) +
  geom_point(size=5, alpha=0.7) +
  geom_text_repel() +
  labs(title = "condition") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  # coord_fixed()
  scale_color_viridis_d(option ="viridis")

gp <- ((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect") & plot_annotation(title = colData(vsd)$experiment[1])
gp %>% print()
}

@@

###### -- Advanced PCA
```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}

for (dds in dds_list) {
vst_dat <- assay(vst(dds))
p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)

# check different PCAs
p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 8,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment

p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 8,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype

p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 8,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition

pairs <- (p1+p2) / (p4) + plot_annotation(title = colData(dds)$experiment[1])
pairs %>% print()
# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

# plot Scree
# find explaining PCs
horn <- parallelPCA(vst_dat)
# horn$n = 4 #aCRY
# warnings()

# horn <- list()
# horn$n <- 4

s <- screeplot(p,
    components = getComponents(p, 1:20),
    vline = c(horn$n, elbow)) +
    geom_label(aes(x = horn$n + 1, y = 30,
      label = paste('Horn',horn$n,sep="="), vjust = -1, size = 8)) +
    geom_label(aes(x = elbow + 1, y = 50,
      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8)) + plot_annotation(title = colData(dds)$experiment[1])

s %>% print()

bi <- biplot(p,x="PC3",y="PC1",
    lab = p$metadata$clientName,
    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    labSize = 3, legendIconSize = 4.0,
    legendPosition = 'bottom',
    sizeLoadingsNames = 3,
    axisLabSize = 10,
    captionLabSize = 1)

@@

layout <- c(
  area(t = 1, l = 1, b = 6, r = 6),
  area(t = 5, l = 1, b = 7, r = 3)
)
pairs_bi <- pairs + bi + 
  plot_layout(design = layout) & plot_annotation(title = colData(dds)$experiment[1])
pairs_bi %>% print()

@@

load <- plotloadings(p,
                     axisLabSize = 8,
             components = getComponents(p)[1:7],
             col = viridis(3),
             labSize = 3) + 
  plot_annotation(title = colData(dds)$experiment[1]) &
  theme(legend.position = 'none', 
        axis.title=element_text(size=8),
        axis.text=element_text(size=6))
load %>% print()
}

@@

## Plot Counts
```{r plot_counts}
dds <- dds_list[[1]]
for (dds in dds_list){
# TF genes

@@

# Plot
gcounts_tf1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_tf1 %>% print()

@@

# Plot
max_val <- 1.0*max(all_counts$count)
gcounts_phot1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_phot1 %>% print()

print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))

}

@@

res_list_list <- list()
for (dds in dds_list) {

@@

@@
